---
output: github_document
editor_options: 
  chunk_output_type: console
---

# Convert WFS data from the Berlin Geodata Portal "FIS-Broker"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(httr)
library(sf)
library(dplyr)
```

```{r}

get_X_Y_coordinates <- function(x) {
  
  sftype <- as.character(sf::st_geometry_type(x, by_geometry = FALSE))

  if(sftype == "POINT") {
    
    xy <- as.data.frame(sf::st_coordinates(x))
    dplyr::bind_cols(x, xy)
    
  } else {
    x
  }
  
}

sf_fisbroker <- function(url) {
  
    typenames <- basename(url)
    
    url <- httr::parse_url(url)
    
    url$query <- list(service = "wfs",
                      version = "2.0.0",
                      request = "GetFeature",
                      srsName = "EPSG:25833",
                      TYPENAMES = typenames)
    
    request <- httr::build_url(url)
    
    print(request)
    
    out <- sf::read_sf(request)
    
    out <- sf::st_transform(out, 4326)
    
    out <- get_X_Y_coordinates(out)

    return(out)
}

export_format <- c(
          "geojson", 
          "sqlite"
   )

sf_save <- function(z, fname) {

  ff <- paste(file.path(fname, fname), export_format, sep = ".")
  purrr::walk(ff, ~{ sf::st_write(z, .x, delete_dsn = TRUE)})
  saveRDS(z, paste0(file.path(fname, fname), ".rds"))
  
}

```

# Description

This repository is a proof-of-concept how to convert WFS data into different output formats using "Simple Features" [https://github.com/r-spatial/sf](https://github.com/r-spatial/sf)

## LOR - Planungsräume

```{r LOR_PLR}

z <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_lor_plan")

z <- z %>%
     mutate(RAUMID = stringr::str_sub(gml_id, 12, 19)) %>%
     select(gml_id, RAUMID, everything()) %>%
     arrange(RAUMID)

dplyr::glimpse(z)

sf_save(z, "LOR_Planungsraum")

```

## Bodenrichtwert 2020

```{r}

z <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_brw_2020")

dplyr::glimpse(z)

sf_save(z, "Bodenrichtwerte_2020")

```

## Gebäudealter der Wohnbebauung 2015

```{r}

z <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s06_12baualter")

dplyr::glimpse(z)

sf_save(z, "Gebaeudealter")

```

## Einwohnerdichte 2018

```{r}

z <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s06_06ewdichte2018")

dplyr::glimpse(z)

sf_save(z, "Einwohnerdichte_2018")

```

## Wohnlagenkarte nach Adressen zum Berliner Mietspiegel 2019

```{r eval = TRUE}

z <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_wohnlagenadr2019")

z  <- z %>% 
       mutate(ADR_num = as.numeric(stringr::str_extract(ADR, "[0-9]+")), 
              ADR_chr = stringr::str_extract(ADR, "[aA-zZ]+"))

dplyr::glimpse(z)

sf_save(z, "Wohnlagenkarte_2019")

```

## Öffentliche Schulen

Data set used in the blog post from [lxndrkp](https://twitter.com/lxndrkp):

[HOW TO: Downloading data from Berlin's geospatial data portal](https://lab.technologiestiftung-berlin.de/projects/fisbroker-to-qgis/index_en.html)

Note: Although the data is published as Open Data I exclude some sensitive data in the export.


```{r eval = TRUE}

z <- sf_fisbroker("https://fbinter.stadt-berlin.de/fb/wfs/data/senstadt/s_schulen") %>%
     select(-TELEFON, -FAX, -EMAIL, -INTERNET)

dplyr::glimpse(z)

sf_save(z, "Schulen")

```
